---
title: "One-Way-ANOVA"
author: "Durjoy Sarker"
format:
  pdf:
    toc: TRUE
    toc-depth: 2
---

# One Way ANOVA


## Load data

```{r}
data <- read.csv("G:/Lesson/Done/Research/R Language/R-Language/Self-Learning/One-Way-ANOVA/fertilizer_yield_for_anova.csv")
data$Fertilizer <- as.factor(data$Fertilizer) # convert to factor so ANOVA treats it as categorical groups, not text or numeric values
```

## Create ANOVA Model

To test whether mean yield differs among fertilizer groups

```{r}
model <- aov(Yield ~ Fertilizer, data = data)
```

## Assumption Checking

Shows one way anova is reliable or not.

### 1. Normality of Residuals

```{r}
shapiro.test(residuals(model))
```

**Interpretation:**

-   p \> 0.05 → residuals are approximately normal → acceptable

-   p \< 0.05 → deviation from normality → check plots before deciding

**Understanding Normality**

-   Normal data → values are evenly distributed around mean. Example: (15, 13, 14, 16)

-   Non-normal data → extreme variation/outliers. Example: (2, 5, 20, 50)

### 2. Homogeneity of Variance

```{r}
library(car)
leveneTest(Yield ~ Fertilizer, data = data)
```

**Interpretation**

1.  P-Value is more than 0.05, the variances across groups are more or less equal (homogeneous), perform the one-way ANOVA reliably.

2.  P-Value is less than 0.05 the variances across groups are not more or less equal (not homogeneous), use caution. see the plot

**Homogeneity**

For understanding homogeneity need to understand variance. Variance means the spread of data in each group.

Example: Group A: (13, 14, 15, 16) → small spread → low variance

Group B: (10, 20, 30, 40) → large spread → high variance

All groups have similar spread, the data is homogeneous. Ad one group is very spread out and others are tight, it's not homogeneous.

### 3. Diagnostic Plots

```{r}
plot(model)
```

**What to check:**

Residuals vs Fitted → should show random pattern (no structure)

Normal Q-Q → points should lie near the straight line

Scale-Location → spread should be consistent

Residuals vs Leverage → detect outliers/influential points

👉 These plots are more reliable than tests. Study more here!!!!

## Perform anova

```{r}
summary(model)
```

**Interpretation:**

p \< 0.05 → significant difference exists among groups

p \> 0.05 → no strong evidence of difference

ANOVA tells “difference exists”. It does NOT tell which groups differ.

## Alternative (Kruskal-Wallis Test)\]

**Use when**

-   strong non-normality

-   outliers present

-   variance highly unequal
```{r}
krus <- read.csv("G:/Lesson/Done/Research/R Language/R-Language/Self-Learning/One-Way-ANOVA/non_normal_yield_data.csv")

kruskal.test(Yield ~ Fertilizer, data = krus)
```

**Interpretation:**

p \< 0.05 → at least one group differs

p \> 0.05 → no evidence of difference











