---
title: "Post_hoc"
author: "Durjoy Sarker"
format:
  pdf:
    toc: TRUE
    toc-depth: 3
---

## ANOVA clean workflow

```{r}
# Load Data & Prepare
data <- read.csv("G:/Lesson/Done/Research/R Language/R-Language/Self-Learning/One-Way-ANOVA/agriculture_posthoc_data.csv")
data$Treatment <- as.factor(data$Treatment)  # ensure categorical variable

#Build ANOVA Model
model <- aov(Yield ~ Treatment, data = data)

#Check Assumptions
##Normality of Residuals
shapiro.test(residuals(model))
```

**Interpretation**

p \> 0.05 → acceptable (not perfect, just no strong violation)

p \< 0.05 → check plots before decision

```{r}
## Homogeneity of Variance

library(car) 
leveneTest(Yield ~ Treatment, data = data)
```

**Interpretation**

p \> 0.05 → equal variance (good)

p \< 0.05 → violation (be careful)

⚠️ Decision Rule (Important)

If assumptions roughly OK → proceed with ANOVA

If strong violation → go to non-parametric (Kruskal)

```{r}
#Run ANOVA 
summary(model)
```

**Interpretation**

p \< 0.05 → reject null → at least one group differs → go for post hoc

p \> 0.05 → fail to reject → stop (no post hoc)

## Normally Distributed data

### 1. Tukey’s HSD (BEST CHOICE)

```{r}
library(agricolae)
tukey_result <- HSD.test(model, "Treatment", group = TRUE) # if group is TRUE, it will show the $group data and $comparison will be null. If the group is FALSE, it will how the comparison data and group will be none
tukey_result
```

**Interpretation**

-   \$means -\> show the mean yield per fertilizer

-   \$comparison -\> signif (stars indicate significance) ..\*\*\* = highly significant (p \< 0.001) if \*\* = p \< 0.01 if \* = p \< 0.05

-   \$Group -\> Same letter= no significant difference, Different letters = significantly different

*Show the specific result:*

`tukey_result$means`

`tukey_result$groups`

`tukey_result$comparison`

**Why best:**

controls error properly accepted in research papers

**Use when:**

-   3+ groups final report / publication

-   Used mainly in agriculture papers

### 2. LSD Test (Least Strict)

```{r}
library(agricolae) 
lsd_result <- LSD.test(model, "Treatment", p.adj = "none")
lsd_result
```

Same interpretation as Tukey.

**LSD limitation:**

LSD test is more liberal (less strict), so it's: Good for few treatments (2–4), exploratory / early analysis. You want high sensitivity (detect small differences)

**Rule of Thumb**

Not good for many groups → higher chance of false positives.

-   2–3 groups → LSD is OK, especially for early or internal analysis

-   3+ groups or final result/report → Prefer Tukey’s HSD

-   If unsure, go with Tukey — it’s safer and accepted

**Use LSD when**

-   2–3 treatments only

-   exploratory / early analysis

-   you want high sensitivity (detect small differences)

**Problem**

high false positives.**That is why go with Tukey.**

### 3. Duncan’s Test (Middle Ground)

```{r}
library(agricolae) 
duncan_result <- duncan.test(model, "Treatment", group = TRUE)
duncan_result
```

**Characteristics**

-   more sensitive than Tukey

-   less strict → more risky

### Alternative using emmeans library(emmeans)

#### LSD equivalent

```{r}
library(emmeans) 
emmeans(model, pairwise ~ Treatment, adjust = "none")
```

#### Tukey

```{r}
library(emmeans) 
emmeans(model, pairwise ~ Treatment, adjust = "tukey") 
```

### Difference Between the Test: 

**1.** **Sensitivity** = How easily the test finds a difference

-   High sensitivity = Finds small differences

-   Low sensitivity = Only finds big differences

**2. Strictness** = How careful the test is to avoid wrong results

-   Strict test = Needs more evidence to say “groups are different”

-   Less strict = Can say “different” even with less evidence (risky)

Final Summary Table:

| Test | When use | Sensitivity | Strictness | Error Control | Most Used |
|----|----|----|----|----|----|
| Turkey | Safe, journal standard, \>3 groups | Lower | High | Strong | Yes |
| LSD | 2-3 groups, quick idea, early research |  High |  Low | None | Less |
| Duncan | Grouping letters, medium strictness | High |  Medium | Some | Some Agri |

## Dunn Test(non-parametric)

If Assumptions FAIL → Non-Parametric Path (Kruskal-Wallis Test)

`kruskal.test(Yield ~ Treatment, data = data)`

**Interpretation**

p \< 0.05 → groups differ → go for **Dunn’s test**

```{r}
library(FSA) 
dunnTest(Yield ~ Treatment, data = data, method = "bonferroni")
```

**Interpretation**

p.adj \< 0.05 → significant difference.
