---
title: "Two-Way-ANOVA"
author: "Durjoy Sarker"
format:
  pdf:
    toc: TRUE
    toc-depth: 3
editor: 
  markdown: 
    wrap: 72
---

## TWO-WAY ANOVA

### Data Loading

```{r}
data <- read.csv("G:/Lesson/Done/Research/R Language/R-Language/Self-Learning/Two_Way_ANOVA/agriculture_two_way_anova.csv")
data$Fertilizer <- as.factor(data$Fertilizer)
data$Irrigation <- as.factor(data$Irrigation)
```

### Model Building

```{r}
model <- aov(Yield ~ Fertilizer * Irrigation, data = data)
```

### Assumption Checking

#### a. Normality of residuals

```{r}
shapiro.test(residuals(model))
```

p \> 0.05 → residuals are normally distributed ✔

#### b. Homogeneity of variance

```{r}
library(car)
leveneTest(Yield ~ Fertilizer * Irrigation, data = data)
```

p \> 0.05 → equal variance assumption satisfied ✔

#### c. Residual diagnostics

```{r}
plot(model)
```

No strong pattern → model is valid ✔

### Perform ANOVA

```{r}
summary(model)
```

**Testing 3 hypotheses:**

-   Fertilizer effect (main effect)

-   Irrigation effect (main effect)

-   Fertilizer × Irrigation interaction (joint effect)

### **Interpretation of ANOVA Table**

**Main Effects-**

Fertilizer effect : p \< 0.05 → Fertilizer types are NOT equal. Means at
least one fertilizer changes yield

Irrigation effect : p \< 0.05 → Irrigation levels significantly affect
yield

**CRITICAL WARNING (Most beginners miss this)**

If interaction is significant, main effects alone are NOT reliable
because:

Effect of fertilizer changes depending on irrigation So “average
fertilizer effect” becomes misleading

**Interaction Effect (Most Important Part)**

Fertilizer × Irrigation: p \< 0.05 → Interaction exists

Correct meaning:

The effect of Fertilizer on Yield depends on Irrigation level (and vice
versa)

NOT:

-   “They are related” ❌

-   “Both are important” ❌

### Proper Statistical Decision Rule

**Case 1: Interaction NOT significant**

Interpret fertilizer main effect and Irrigation main effect separately

**Case 2: Interaction significant**

IGNORE main effects

Focus on: Interaction plots, Post-hoc within each irrigation level

### Interaction Plot

```{r}
with(data, interaction.plot(Fertilizer, Irrigation, Yield))
```

#### **Interpret Interaction Plot:**

**Step 1: Check line crossing**

-   Lines crossing → strong interaction

-   Lines parallel → no interaction

**Step 2: Interpret patterns scientifically**

Example interpretation:

**Fertilizer A:**

High irrigation → highest yield

Low irrigation → sharp drop

That's indicate **HIGH water dependency**

**Fertilizer B:**

Stable across irrigation levels → drought-tolerant performance

**Fertilizer C:**

Low performance everywhere → generally inefficient

8.  Stronger Statistical Add-ons (YOU SHOULD ADD THIS) (A) Mean Table
    (Very Important in research) aggregate(Yield \~ Fertilizer +
    Irrigation, data = data, mean) (B) Post-hoc Test (ONLY if needed)

If interaction is significant:

TukeyHSD(model)

But better approach:

Analyze simple effects instead of global Tukey blindly
