---
title: "Significance test"
author: "Durjoy Sarker"
format:
  pdf:
    toc: TRUE
    toc-depth: 2 
---

# ttest

```{r}
roll <- c(23,45,66,34,54) 
marks <- c(87,89,80,98,87) 
student_data <- data.frame(roll,marks) 
student_data$group <- ifelse(student_data$roll < 50, "Group1", "Group2") 
table(student_data$group) 
ttest <- t.test(marks ~ group,data = student_data) 
ttest  
```

# One Way ANOVA

```{r}
Stu_data <- data.frame( 
Method = rep(c("A", "B", "C"), each = 5), 
Score = c(85, 90, 88, 87, 84,  # Method A 
78, 76, 80, 75, 79,  # Method B 
92, 94, 95, 91, 93)  # Method C 
) 

anova_result <- aov(Score ~ Method, data = Stu_data) 
summary(anova_result) 
```

-   aov: A function in R for conducting ANOVA.

-   Score: The dependent variable or the response variable.

-   Method: The independent variable or the factor whose levels you are

**Look at the p-value in the output:**

1.  If p-value \< 0.05, reject the null hypothesis (there is a significant difference between the groups).

2.  If p-value ≥ 0.05, fail to reject the null hypothesis (there is no significant difference).

*The null hypothesis (H₀) is the default assumption that there is no difference between the groups.*
