---
title: "Two-Way-ANOVA"
author: "Durjoy Sarker"
format:
  pdf:
    toc: TRUE
    toc-depth: 3
editor: 
  markdown: 
    wrap: 72
---

## TWO-WAY ANOVA

### Data Loading

```{r}
data <- read.csv("G:/Lesson/Done/Research/R Language/R-Language/Self-Learning/Two_Way_ANOVA/agriculture_two_way_anova.csv")
data$Fertilizer <- as.factor(data$Fertilizer)
data$Irrigation <- as.factor(data$Irrigation)
```

### Model Building

```{r}
model <- aov(Yield ~ Fertilizer * Irrigation, data = data)
```

### Assumption Checking

#### a. Normality of residuals

```{r}
shapiro.test(residuals(model))
```

p \> 0.05 → residuals are normally distributed ✔

#### b. Homogeneity of variance

```{r}
library(car)
leveneTest(Yield ~ Fertilizer * Irrigation, data = data)
```

p \> 0.05 → equal variance assumption satisfied ✔

#### c. Residual diagnostics

```{r}
plot(model)
```

No strong pattern → model is valid ✔

### Perform ANOVA

```{r}
summary(model)
```

**Testing 3 hypotheses:**

-   Fertilizer effect (main effect)

-   Irrigation effect (main effect)

-   Fertilizer × Irrigation interaction (joint effect)

### Interpretation of ANOVA Table

**Main Effects-**

Fertilizer effect : p \< 0.05 → Fertilizer types are NOT equal. Means at
least one fertilizer changes yield

Irrigation effect : p \< 0.05 → Irrigation levels significantly affect
yield

**CRITICAL WARNING (Most beginners miss this)**

If interaction is significant, main effects alone are NOT reliable
because:

Effect of fertilizer changes depending on irrigation So “average
fertilizer effect” becomes misleading

**Interaction Effect (Most Important Part)**

Fertilizer × Irrigation: p \< 0.05 → Interaction exists

Correct meaning:

The effect of Fertilizer on Yield depends on Irrigation level (and vice
versa)

NOT:

-   “They are related” ❌

-   “Both are important” ❌

### Proper Statistical Decision Rule

**Case 1: Interaction NOT significant**

Interpret fertilizer main effect and Irrigation main effect separately

**Case 2: Interaction significant**

IGNORE main effects

Focus on: Interaction plots, Post-hoc within each irrigation level

### Interaction Plot

```{r}
with(data, interaction.plot(Fertilizer, Irrigation, Yield))
```

#### Interpret Interaction Plot:

**Step 1: Check line crossing**

-   Lines crossing → strong interaction

-   Lines parallel → no interaction

**Step 2: Interpret patterns scientifically**

Example interpretation:

**Fertilizer A:**

High irrigation → highest yield

Low irrigation → sharp drop

That's indicate **HIGH water dependency**

**Fertilizer B:**

Stable across irrigation levels → drought-tolerant performance

**Fertilizer C:**

Low performance everywhere → generally inefficient

### Mean table

```{r}
aggregate(Yield ~ Fertilizer + Irrigation, data = data, mean)
```

## POST HOC
### Global turkey
```{r}
TukeyHSD(model)
```

**Problem with Global Tukey**

This compares all combinations together, which:

-   mixes irrigation + fertilizer effects

-   hides the actual interaction structure

👉 Result = confusing + misleading conclusions. That's why this is
technically correct but scientifically wrong.

### Correct Approach: Simple Effects Analysis
#### 1. Effect of Fertilizer at each Irrigation level
```{r}
#Effects of fertilizer when irrigation low
model_low <- aov(Yield ~ Fertilizer, data = subset(data, Irrigation == "Low"))
TukeyHSD(model_low)


#Effects of fertilizer when irrigation high
model_high <- aov(Yield ~ Fertilizer, data = subset(data, Irrigation == "High"))
TukeyHSD(model_high)
```

#### 2. Effect of Irrigation at each Fertilizer
```{r}
model_A <- aov(Yield ~ Irrigation, data = subset(data, Fertilizer == "A"))
TukeyHSD(model_A)
```





